A Journey from Wild to Textbook Data to Reproducibly Refresh the Wages Data from the National Longitudinal Survey of Youth Database

### 2022-02-16 paper/paper.Rmd
filtering dropouts

An additional criteria from @SingerJudithD2003Alda is to only include males aged between 14 and 17 years old in 1979. Filters using these criteria used re-create the dropouts data.

With this filtering, we obtained 670 individuals in the refreshed data compared to 888 individuals in the original data. To investigate the reason for the difference, individuals from the original and refreshed dataset were matched by id. This revealed several reasons for for the disparity:

1. 173 individuals were more than 17 years old in 1979. Thus, it looks like the description of the original data is not quite accurate, that there are people older than 17 in the subset. Our decision is to also include them in the refreshed data as the new data contains an age variable, so analyst could filter them later. 
2. 79 individuals were less than or equal to 17 years old in 1979. However, they were not captured in the refreshed data because:

  + 35 of them completed at least 12th grade with a diploma instead of GED (`ged` variable is coded to 1). This suggests that they are not dropouts, and so we excluded them from the refreshed data. 
  + The information about `ged` is missing in 38 individuals. We decided to include them in the refreshed data. 
  + 3 individuals have both diploma and GED (`ged` is coded to 3). These were kept in the refreshed data. 
  + 12 individuals do not exist in `wages` data because they were participated in less than 3 rounds of survey.

The filtering was re-applied using these decisions. The refreshed dropout data has 863 individuals.

>>>>

An additional criteria from @SingerJudithD2003Alda is to only include males aged between 14 and 17 years old in 1979. With this filtering, we obtained 670 individuals in the refreshed data compared to 888 individuals in the original data. To investigate the reason for the difference, individuals from the original and refreshed dataset were matched by id. This revealed several reasons for for the disparity:

1. 173 individuals were more than 17 years old in 1979. Thus, it looks like the description of the original data is not quite accurate, that there are people older than 17 in the subset. Our decision is to also include them in the refreshed data as the new data contains an age variable, so analyst could filter them later. 
2. 79 individuals were less than or equal to 17 years old in 1979. However, they were not captured in the refreshed data because: 
    i. 35 of them completed at least 12th grade with a diploma instead of GED (`ged` variable is coded to 1). This suggests that they are not dropouts, and so we excluded them from the refreshed data. 
    ii. The information about `ged` is missing in 38 individuals. We decided to include them in the refreshed data. 
    iii. 3 individuals have both diploma and GED (`ged` is coded to 3). These were kept in the refreshed data. 
    iv. 12 individuals do not exist in `wages` data because they were participated in less than 3 rounds of survey.

The filtering was re-applied using these decisions, resulting in a refreshed dropout subset containing 863 individuals.

### 2022-02-16 paper/paper.tex
filtering dropouts

\begin{verbatim}
#> # A tibble: 2 x 3
#>   hgc   hgc12     n
#>   <chr> <chr> <int>
#> 1 12TH  8TH       1
#> 2 8TH   12TH    243
\end{verbatim}

@@

An additional criteria from Singer and Willett (2003) is to only include males aged between 14 and 17 years old in 1979. Filters using these criteria used re-create the dropouts data.

With this filtering, we obtained 670 individuals in the refreshed data compared to 888 individuals in the original data. To investigate the reason for the difference, individuals from the original and refreshed dataset were matched by id. This revealed several reasons for for the disparity:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  173 individuals were more than 17 years old in 1979. Thus, it looks like the description of the original data is not quite accurate, that there are people older than 17 in the subset. Our decision is to also include them in the refreshed data as the new data contains an age variable, so analyst could filter them later.
\item
  79 individuals were less than or equal to 17 years old in 1979. However, they were not captured in the refreshed data because:
\end{enumerate}

\begin{itemize}
\tightlist
\item
  35 of them completed at least 12th grade with a diploma instead of GED (\texttt{ged} variable is coded to 1). This suggests that they are not dropouts, and so we excluded them from the refreshed data.
\item
  The information about \texttt{ged} is missing in 38 individuals. We decided to include them in the refreshed data.
\item
  3 individuals have both diploma and GED (\texttt{ged} is coded to 3). These were kept in the refreshed data.
\item
  12 individuals do not exist in \texttt{wages} data because they were participated in less than 3 rounds of survey.
\end{itemize}

The filtering was re-applied using these decisions. The refreshed dropout data has 863 individuals.

>>>>

@@

An additional criteria from Singer and Willett (2003) is to only include males aged between 14 and 17 years old in 1979. With this filtering, we obtained 670 individuals in the refreshed data compared to 888 individuals in the original data. To investigate the reason for the difference, individuals from the original and refreshed dataset were matched by id. This revealed several reasons for for the disparity:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  173 individuals were more than 17 years old in 1979. Thus, it looks like the description of the original data is not quite accurate, that there are people older than 17 in the subset. Our decision is to also include them in the refreshed data as the new data contains an age variable, so analyst could filter them later.
\item
  79 individuals were less than or equal to 17 years old in 1979. However, they were not captured in the refreshed data because:

  \begin{enumerate}
  \def\labelenumii{\roman{enumii}.}
  \tightlist
  \item
    35 of them completed at least 12th grade with a diploma instead of GED (\texttt{ged} variable is coded to 1). This suggests that they are not dropouts, and so we excluded them from the refreshed data.
  \item
    The information about \texttt{ged} is missing in 38 individuals. We decided to include them in the refreshed data.
  \item
    3 individuals have both diploma and GED (\texttt{ged} is coded to 3). These were kept in the refreshed data.
  \item
    12 individuals do not exist in \texttt{wages} data because they were participated in less than 3 rounds of survey.
  \end{enumerate}
\end{enumerate}

The filtering was re-applied using these decisions, resulting in a refreshed dropout subset containing 863 individuals.
